\hypertarget{problems}{%
\section{Problems}\label{problems}}

\begin{enumerate}
\tightlist
\item
  An object is displaced by \(<10,5,15>\) and then by \(<-5, 20, 0>\).
  What is the total displacement?
\item
  What is the angle between \(<1,2,3>\) and \(<3,2,1>\)?
\item
  Given (1,2), (2,1), (3,2), (4,4).

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Using the pseudoinverse, find the least squares linear interpolant
    of the data points: \(y = a_1x + a_0\).
  \item
    Plot the data and the resulting line.
  \end{enumerate}
\item
  Given \(x+y = 1\), \(x+2y = 0\), \(2x+y = 1\).

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Write this as a linear system and find the left pseudoinverse of the
    matrix. {[}Note: label the matrix as A, right hand side as b and
    unknown as x.{]}
  \item
    Find the least squares solution for \((x_1, x_2)\).
  \item
    What is the error \(\| Ax - b\|\)?
  \end{enumerate}
\item
  Given \(x+2y = 2\), \(2x+4y = 4\).

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    Frame this as a 2x2 matrix problem.
  \item
    Can you solve this system? Why or why not?
  \item
    Plot the columns of the matrix and the right hand side (as vectors).
  \item
    Plot \(A^Tb\).
  \item
    What is the least squares solution for x?
  \end{enumerate}
\item
  Find the 3x3 rotation matrix that will rotate

  \begin{enumerate}
  \def\labelenumii{\alph{enumii}.}
  \tightlist
  \item
    about the x axis by \(30^\circ\).
  \item
    about the z axis by \(120^\circ\).
  \end{enumerate}
\item
  Analytically show that the inverse rotation matrix is the same matrix
  as replacing \(\theta\) by \(-\theta\).
\item
  In homogeneous coordinates, is a rotation followed by a displacement
  the same as the displacement followed by the rotation? Prove if this
  is true or provide a counter-example. Note that proof does not mean
  give examples.
\item
  Find the 3x3 transformation matrix that rotates a vector about y by
  \(15^\circ\) and then about x by \(25^\circ\).
\item
  Find the 4x4 homogeneous coordinate transformation matrix that rotates
  a vector about z by \(45^\circ\), then translates by \(<1,2,3>\) and
  then rotates about y by \(30^\circ\).
\item
  Show that one of the eigenvectors for a rotation matrix points in the
  direction of the axis of rotation. What is the eigenvalue?
\item
  Assume that you have a wheel rolling on a circular path (or track).
  Let the radius of the path be 10m, the radius of the wheel be 30cm.
  Assume the wheel is turning at 3 revolutions per minute. What is the
  revolutions per minute around the path?
\end{enumerate}
